---
title: "Frequency of solidago altissima in North American and Canadian regions."
name: "Tyler Edvall"
date: "03/11/21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Abstract 

Solidago altissima is a Northern American species of goldenrod plant. They are often found across large areas of the United States, northern parts of Mexico, and Canada. Solidago altissima is unique due to its tolerance of landscapes which have had human disturbance. This unique trait has allowed the plant to grow widespread in the North American/Canadian region. The question I will answer explores if there is a higher frequency of solidago altissima plants in areas of the United States compared to Canada. Data being used for this question comes from the “BIEN” database, which is the “Botanical Information and Ecology Network”. Within this database, there is an option to search for worldwide data observations on a specific plant species. I will use exploratory data analysis to answer my question. The two variables that will be compared in the dataset are the observations, and the country. I will separate the observations variable by “specimen” and the country variable by “United States” and “Canada”. The purpose of this research question would relate to ecological conservation and in what human disturbed areas solidago altissima growth frequency is higher. As well as, studying how plants can better adapt to the rapidly growing human disturbed areas. I predict that the frequency of solidago altissima will be higher in the region of Canada, due to less human disturbed areas. 

#Introduction

Solidago altissima is a Northern American species of goldenrod plant. They are often found across large areas of the United States, northern parts of Mexico, and Canada. Their native habitat includes dry to moist roadsides, thickets, prairies & open woods, making Solidago altissima common in human populated areas. Solidago altissima is unique due to its tolerance of landscapes which have had human disturbance. Another unique trait is the plant's high plasticity to various soil types. Solidago altissima is able to grow in moist to dry soils, clay, clay loam, medium loam, sandy loam, Sandy, and caliche type soils. These unique traits have allowed the plant to grow widespread in the North American/Canadian region.


#Methods



##Data Acquisition

*Solidago altissima data was collected from the Botanical Information and Ecology Network (BIEN). 
*BIEN version used was 4.1.1, 2018-12-06. 
*Packages used:

```{r}

citation()
citation("BIEN")
citation("dplyr")
citation("ggplot2")


```


##Data Preparation

*Graphed data was filtered from the library(BIEN) function.
*library(ape) package was used for working with phylogenies in R.
*library(maps) used for making quick maps of occurrences.
*library(sp) the package was used for spatial data.
*BIEN data was filtered by species, species occurrence, new world only, country, United States, Canada.
*Solidago altissima frequency was counted in both United States and Canada. 



#Results



#Discussion



#Literature Cited



#References 